---
title: "Root_Fungi"
author: "Wes Bickford"
date: "2/16/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(permute); library(vegan); library("dplyr"); library(knitr)
source("data_setup.R")
```
#Root Fungi Analysis

```{r Rarefaction setup, echo = F, include = F}
source('data_setup.R')
mean(metadata$n_seqs)

rarefy <- read.table(file = "data/Bickford.groups.rarefaction", header=T, stringsAsFactors=F)

ave_columns <- grep(pattern = "X1", colnames(rarefy))
ave_rarefy <- rarefy[ , ave_columns]

numsampled <- rarefy$numsampled

sample_IDs <- gsub(pattern = "X1.", replacement = "", x=colnames(ave_rarefy))
colnames(ave_rarefy) <- sample_IDs

clrs <- c(BL = "grey", CB = "blue", CH = "red", CM = "coral", CR = "black", PLB = "orange", Rt2 = "green", SB = "purple")


```

##Rarefaction Curves

These are rarefaction curves for all samples. Samples were rarefied to 200 sequences as that was the lowest number at any site. 

```{r Rarefaction plot, echo = F, include = T}
plot(NA, type = "l",
     xlab = "Number of Sequences Sampled",
     ylab = "Number of OTUs Observed",
     ylim = c(0,150),
     xlim = c(0,3000)
     )
for(i in sample_IDs){
      points(ave_rarefy[ ,i] ~ numsampled, type= "l",
             col=clrs[metadata[metadata$sample==i, "Site"]],
             lwd = 2, lty = 1
       )
}
abline(v=200)
```

##Alpha Diversity 
```{r alpha setup, echo = F, include = F}
alpha <- read.table(file = "data/Bickford.groups.ave-std.summary", header=T)

alpha_mean <- alpha[alpha$method == "ave", ]
alpha_mean$group <- as.character(alpha_mean$group)

meta_alpha <- inner_join(metadata, alpha_mean, by = c("sampleID"="group"))

# removing PLB and SB from this plot because there are no replicates
meta_alpha_sub <- meta_alpha[meta_alpha$Site != "SB" & meta_alpha$Site != "PLB", ]
meta_alpha_sub$Site <- as.character(meta_alpha_sub$Site)
meta_alpha_sub$Site <- as.factor(meta_alpha_sub$Site)
```

The following shows shannon diversity by site and lineage

```{r alpha plots, echo = F, include = T}
clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", Rt2 = "gray")

boxplot(npshannon ~ Site + Lineage, data = meta_alpha_sub,
        ylab = "Shannon Diversity Index",
        ylim = c(0,5.5), col = clrs[levels(meta_alpha_sub$Site)],
        names = rep(x = unique(meta_alpha_sub$Site), times = 2),
        range = 0)

Lineage_convert <- c(Nat = "Native", Inv = "Non-Native")
mtext(side=1, at=c(3.5,9.5), line=3, text = Lineage_convert[levels(meta_alpha$Lineage)], cex = 1.5, font = 2)

```

Histograms show that shannon diversity is fairly normally distributed.

```{r shannon hists, echo = F, include=T}
hist((meta_alpha_sub$npshannon), breaks = 6)
hist(meta_alpha$npshannon, breaks = 6)
```

Assuming those are normally distributed, here are the results of an ANOVA by site and lineage. No significant difference in alpha diversity by site or lineage. 

```{r ANOVA shannon, echo = F, include = T}
an_shannon <- aov(npshannon ~ Site * Lineage, data = meta_alpha)
summary(an_shannon)
```

##Community Analysis
###PerManova

The following is a PerMANOVA of phyloytpes by site and Lineage

```{r permanova1, include=T, echo = F, results = 'hold'}
shared <- read.table(file= "data/Bickford.subsample.shared", header = T, stringsAsFactors = F, row.names = 2)
shared <- shared[ ,-c(1,2)]

permfungi <- adonis(shared ~ Site * Lineage, data=metadata, method = "bray")
permfungi

```


There is a significant *site* effect. Lineage and their interaction are non-significant.

###NMDS

```{r NMDS, include = F, echo = F}

NMDS <- metaMDS(shared, try = 20, trymax = 300, k=2)
NMDS_scores <- scores(NMDS)

NMDS1 <- NMDS_scores[ ,1]
NMDS2 <- NMDS_scores[ ,2]

NMDS_scores <- as.data.frame(NMDS_scores)
```

The following is an NMDS (stress = `r NMDS$stress`) showing Lineage as different shapes and site by color. The ellipses are sd by site. 

```{r NMDS plot, include=T, echo = F}

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1,1.8), ylim = c(-1,1))
axis(side = 1)
axis(side = 2, las = 1)
points(x = NMDS_scores$NMDS1, y = NMDS_scores$NMDS2, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1.5)

mtext(side = 1, line = 2, text = "NMDS1")
mtext(side = 2, line = 3, text = "NMDS2" )

ordiellipse(NMDS, metadata$Site == "BL", show.groups = TRUE, kind = "sd", col = "blue", lty = 1, lwd = 2.0)
ordiellipse(NMDS, metadata$Site == "CB", show.groups = TRUE, kind = "sd", col = "red", lty = 1, lwd = 2.0)
ordiellipse(NMDS, metadata$Site == "CH", show.groups = TRUE, kind = "sd", col = "dark green", lty = 1, lwd = 2.0)
ordiellipse(NMDS, metadata$Site == "CM", show.groups = TRUE, kind = "sd", col = "coral", lty = 1, lwd = 2.0)
ordiellipse(NMDS, metadata$Site == "CR", show.groups = TRUE, kind = "sd", col = "black", lty = 1, lwd = 2.0)
ordiellipse(NMDS, metadata$Site == "Rt2", show.groups = TRUE, kind = "sd", col = "gray", lty = 1, lwd = 2.0)

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("right", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```

###PCA

```{r PCA plot, echo = F, include = T}
pca <- read.table(file="data/Bickford.subsample.1.pca.axes", header=T)

#this makes sure the samples are in the same order in both files
stopifnot(pca$group == metadata$sampleID)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
sym <- c(Inv = 23, Nat = 21)

plot(x=pca$axis1, y=pca$axis2, xlab="PCA Axis 1", ylab="PCA Axis 2",
     xlim=c(-1, max(pca$axis1)), ylim=c(min(pca$axis2), max(pca$axis2)), 
     pch=sym[as.character(metadata$Lineage)], col="black", bg= clrs[as.character(metadata$Site)], cex=1, lwd = 1,
     main="PCA by Site and Lineage")

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("topleft", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("top", legend_text2, pch = c(1,5))
```

###Remove rare taxa

I removed rare taxa using relative abundance data. Any OTU that made up less than 0.1% of the total sequences was removed. This is more stringent than removing singletons, but ensures that rare taxa are not responsible for site effects. The following are results of a permanova of only abundant taxa by site and lineage. 

```{r, Permanova abundant, echo = F, include = T}
shared <- read.table(file="data/Bickford.subsample.shared", header=T, stringsAsFactors=F, row.names=2)
shared <- shared[,-c(1,2)]

rel_abund <- shared / apply(shared, 1, sum)
mean_rel_abund <- apply(rel_abund, 2, mean)
otu_order <- order(mean_rel_abund, decreasing=T)
rel_abund <- rel_abund[,otu_order]

stopifnot(rownames(shared) == metadata$sampleID)

mean_rel_abund <- apply(rel_abund, 2, mean)
abundant <- mean_rel_abund > 0.001
rel_abund_subset <- rel_abund[, abundant]
shared_abundant <- shared[ , abundant]

permabundant <- adonis(shared_abundant ~ Site * Lineage, data = metadata)
permabundant
```

Removing rare taxa reduced the total OTU number from `r length(colnames(shared))` to `r length(colnames(shared_abundant))`

There is still a significant difference in the community using only abundant taxa. 

I also performed a Kruskal-Wallace test to see if the relative abundance of any individual taxa differed by site or lineage. All tests were non-significant indicating that no individual otu relative abundances are significantly different by site or lineage 


##Taxonomic Analysis

```{r Taxonomy setup, echo = F, include = F}
taxonomy <- read.table(file = "data/Bickford.cons.taxonomy", header=T, stringsAsFactors = F)


funtax <- taxonomy$Taxonomy[grep(pattern = "Fungi", x=taxonomy$Taxonomy)]
plant_otu <- taxonomy[grep(pattern = "Plant", x=taxonomy$Taxonomy), "OTU"]
protist_otu <- taxonomy[grep(pattern = "Protis", x=taxonomy$Taxonomy), "OTU"]

# this separates by phylum
taxnocon <- gsub(pattern = "\\(\\d*\\)", replacement = "", x=taxonomy$Taxonomy)
nofungi <- gsub(pattern = "k__Fungi;", replacement = "", x=taxnocon)
nobegin <- gsub(pattern = "p__", replacement = "", x=nofungi)
phylum <- gsub(pattern = ";.*", replacement = "", x=nobegin)

# by genus
uptogenus <- gsub(pattern = "k_.*g__", replacement = "", x=taxnocon)
addunclass <- gsub(pattern = "k_.*;unclassified.*", replacement = "unclassified", x=uptogenus)
moreunclass <- gsub(pattern = "unclassified_.*", replacement = "unclassified", x=addunclass)
genus <- gsub(pattern = ";.*", replacement = "", x=moreunclass)
  
# i want a column with otu, phylum, and genus

otu_phylum_genus <- data.frame(otu = taxonomy$OTU, phylum = phylum, genus = genus, stringsAsFactors = F)

shared <- read.table(file= "data/Bickford.subsample.shared", header = T, stringsAsFactors = F, row.names = 2)
shared <- shared[ ,-c(1,2)]


L <- otu_phylum_genus$otu %in% colnames(shared)
otu_phylum_genus <- otu_phylum_genus[L, ]
stopifnot(nrow(otu_phylum_genus)==ncol(shared))

# how do I get the otus from the Ascomycota?

asco_otus <- otu_phylum_genus[otu_phylum_genus$phylum == "Ascomycota", "otu"]
asco_shared <- shared[ , asco_otus]

# how do I combine the number of sequences from each of the Acomycota OTUs

    # in the below, 1 is for rows and 2 is for columns
ascomycota_count <- apply(asco_shared, 1, sum)

# how do I repeat this for all phyla?
#make a loop

#need to know the phylum names - phylum_names / phylum_name
phylum_names <- unique(otu_phylum_genus$phylum)
# need somwhere to put data - allphylum_shared
n_phyla <- length(phylum_names)
n_samples <- nrow(shared)
allphylum_shared <- data.frame(matrix(0, nrow = n_samples, ncol = n_phyla))
rownames(allphylum_shared) <- rownames(shared)
colnames(allphylum_shared) <- phylum_names

#my loop
for(i in phylum_names){
  phylum_otus <- otu_phylum_genus[otu_phylum_genus$phylum == i, "otu"]
  phylum_shared <- shared[ , phylum_otus]
  if(length(phylum_otus) > 1){
    phylum_count <- apply(phylum_shared, 1, sum)
    } else {
      phylum_count <- phylum_shared
    }
  allphylum_shared[ , i] <- phylum_count
  
}

# 3. calculate the relative abundance
n_seqs <- apply(allphylum_shared, 1, sum)
phylum_rel_abund <- allphylum_shared / n_seqs[1]

tot_rel_abund <- numeric()
for(h in colnames(phylum_rel_abund)){
  y <- mean(phylum_rel_abund[ ,h])
  tot_rel_abund[h] <- y
}
sum(tot_rel_abund)
```
###By Phylum

The following show all fungal Phyla by relative sequence abundance

```{r Phylum relative abundance, echo = F, include = T}
shorter_names <- tot_rel_abund
names(shorter_names) <- c("Asco", "Unclass", "Unclass", "Basidio", "Zygo", "Rozello", "Glomero")
barplot(height = shorter_names, beside=T)

```

Ascomycota makes up `r (tot_rel_abund['Ascomycota'])*100`% of the total sequences!

Glomeromycota makes up only `r (tot_rel_abund['Glomeromycota'])*100`%!!

###By Genus

```{r Rel abund setup (genus), echo = F, include = F}
otu_asco_genus <- otu_phylum_genus[otu_phylum_genus$phylum == "Ascomycota", ]

asco_otus <- otu_phylum_genus[otu_phylum_genus$phylum == "Ascomycota", "otu"]
asco_shared <- shared[ , asco_otus]

# how do I combine the number of sequences from each of the Acomycota OTUs

# in the below, 1 is for rows and 2 is for columns
ascomycota_count <- apply(asco_shared, 1, sum)

# how do I repeat this for all phyla?
#make a loop

#need to know the genus names within ascomycota
genus_names <- unique(otu_asco_genus$genus)
# need somwhere to put data - asco_genera_shared
n_genera <- length(genus_names)
n_samples <- nrow(asco_shared)
asco_genera_shared <- data.frame(matrix(0, nrow = n_samples, ncol = n_genera))
rownames(asco_genera_shared) <- rownames(asco_shared)
colnames(asco_genera_shared) <- genus_names

#my loop
for(i in genus_names){
  genus_otus <- otu_asco_genus[otu_asco_genus$genus == i, "otu"]
  genus_shared <- shared[ , genus_otus]
  if(length(genus_otus) > 1){
    genus_count <- apply(genus_shared, 1, sum)
  } else {
    genus_count <- genus_shared
  }
  asco_genera_shared[ , i] <- genus_count
  
}


# 3. calculate the relative abundance
n_seqs <- apply(asco_genera_shared, 1, sum)
genus_rel_abund <- asco_genera_shared / n_seqs

tot_genus_rel_abund <- numeric()
for(h in colnames(genus_rel_abund)){
  y <- mean(genus_rel_abund[ ,h])
  tot_genus_rel_abund[h] <- y
}
sum(tot_genus_rel_abund)
tot_genus_rel_abund

L <- sort.list(tot_genus_rel_abund, decreasing = T)
tot_genus_rel_abund <- tot_genus_rel_abund[L]

genus_names <- unique(otu_asco_genus$genus)
cl_genus_names <- genus_names[genus_names != "unclassified"]
# need somwhere to put data - asco_genera_shared
n_genera <- length(cl_genus_names)
n_samples <- nrow(asco_shared)
asco_cl_genera_shared <- data.frame(matrix(0, nrow = n_samples, ncol = n_genera))
rownames(asco_cl_genera_shared) <- rownames(asco_shared)
colnames(asco_cl_genera_shared) <- cl_genus_names

#my loop
for(i in cl_genus_names){
  genus_otus <- otu_asco_genus[otu_asco_genus$genus == i, "otu"]
  genus_shared <- shared[ , genus_otus]
  if(length(genus_otus) > 1){
    genus_count <- apply(genus_shared, 1, sum)
  } else {
    genus_count <- genus_shared
  }
  asco_cl_genera_shared[ , i] <- genus_count
  
}

# 3. calculate the relative abundance
n_seqs <- apply(asco_cl_genera_shared, 1, sum)
cl_genus_rel_abund <- asco_cl_genera_shared / n_seqs

tot_cl_genus_rel_abund <- numeric()
for(h in colnames(cl_genus_rel_abund)){
  y <- mean(cl_genus_rel_abund[ ,h])
  tot_cl_genus_rel_abund[h] <- y
}
sum(tot_cl_genus_rel_abund)
tot_cl_genus_rel_abund

L <- sort.list(tot_cl_genus_rel_abund, decreasing = T)
tot_cl_genus_rel_abund <- tot_cl_genus_rel_abund[L]

top_genera_rel_abund <- tot_cl_genus_rel_abund[tot_cl_genus_rel_abund >= 0.01]

```

The following is a plot of relative abundance by genus within Ascomycota. I have removed all genera that were unclassified. Unclassified genera made up `r tot_genus_rel_abund['unclassified']*100`% of the total Ascomycota.  
Only the genera with a relative abundance more than 1% of classified genera are displayed here. 

```{r Rel abund by Genus plot, echo=F, include = T}
top_genera_rel_abund

short_genus <- top_genera_rel_abund
names(short_genus) <- c("G", "T", "Mi", "C", "St", "I", "Sa", "F", "As", "N", "Ac", "My", "P", "Mo", "Al")
barplot(height = short_genus, beside=T)


```

### By Function
I ran my taxonomy data through the FUNGuild database to get functional guilds for all taxa. Unclassified taxa were obviously not matched. FUNGuild gives trophic mode and Guild as outputs. Guild is nested within trophic mode, so I have shown the data by trophic mode, then guild

#### Trophic Mode

```{r Funguild setup, echo = F, include = F}
guilds <- read.table("data/for_funguild.guilds_matched.txt", header = T, sep = "\t", stringsAsFactors = F)

#make metadata file for guilds
guild_data <- guilds[ , -c(2:48)]
guild_data$Trophic.Mode <- as.factor(guild_data$Trophic.Mode)
guild_data$Guild <- as.factor(guild_data$Guild)
summary(guild_data)

matched_guild_shared <- guilds[ , c(1:48)]
row.names(matched_guild_shared) <- matched_guild_shared$OTU
matched_guild_shared <- matched_guild_shared[ , -1]

### I want rel abundance of each guild
trans_matched_guild_shared <- data.frame(t(matched_guild_shared))

# how do I combine the number of sequences from each of the OTUs

# in the below, 1 is for rows and 2 is for columns
matched_guild_count <- apply(trans_matched_guild_shared, 1, sum)

# how do I repeat this for all phyla?
#make a loop

#need to know the unique trophic modes 
trophic_names <- unique(guild_data$Trophic.Mode)
# need somwhere to put data - a new shared file
n_modes <- length(trophic_names)
n_samples <- nrow(trans_matched_guild_shared)
trophic_shared <- data.frame(matrix(0, nrow = n_samples, ncol = n_modes))
rownames(trophic_shared) <- rownames(trans_matched_guild_shared)
colnames(trophic_shared) <- trophic_names

#my loop
for(i in trophic_names){
  trophic_otus <- guild_data[guild_data$Trophic.Mode == i, "OTU"]
  guild_shared <- shared[ , trophic_otus]
  if(length(trophic_otus) > 1){
    trophic_count <- apply(guild_shared, 1, sum)
  } else {
    trophic_count <- guild_shared
  }
  trophic_shared[ , i] <- trophic_count
  
}

# 3. calculate the relative abundance
n_seqs <- apply(trophic_shared, 1, sum)
trophic_rel_abund <- trophic_shared / n_seqs

tot_trophic_rel_abund <- numeric()
for(h in colnames(trophic_rel_abund)){
  y <- mean(trophic_rel_abund[ ,h])
  tot_trophic_rel_abund[h] <- y
}
sum(tot_trophic_rel_abund)
tot_trophic_rel_abund

L <- sort.list(tot_trophic_rel_abund, decreasing = T)
tot_trophic_rel_abund <- tot_trophic_rel_abund[L]


```

I created a plot of relative abundance by trophic mode

```{r Trophic mode plot, echo = F, include = T}
tot_trophic_rel_abund
short_trophic <- tot_trophic_rel_abund
names(short_trophic)<- c("Path", "Sap", "P-Sym", "Sap-P", "Sym", "P-Sap-Sym")
barplot(height = short_trophic, beside=T)

```

Pathogens are most abundant. The pathotroph-symbiotroph category is difficult to know what to do with. I assume they have been shown to have multiple trophic modes depending on the host.  

#####Community by Trophic mode
Here is a community analysis by trophic mode. First I ran a Permanova to see if the sites had significatly different communities of trophic modes.

```{r perm by trophic mode, echo = F, include = T}
perm_trophic <- adonis(trophic_shared ~ Site * Lineage, data = metadata)
perm_trophic

```

```{r trophic NMDS setup, echo = F, include = F}
trophicNMDS <- metaMDS(trophic_shared, try = 20, trymax = 300, k=2)
trophic_scores <- scores(trophicNMDS)

trophic_scores <- as.data.frame(trophic_scores)

```

Here is an NMDS grouped by trophic mode (stress = `r trophicNMDS$stress`)

```{r trophicNMDS plot, echo = F, include = T}
quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1,1.8), ylim = c(-1,1))
axis(side = 1)
axis(side = 2, las = 1)
clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
points(x = trophic_scores$NMDS1, y = trophic_scores$NMDS2, pch = pchs[as.character(metadata$Lineage)], 
       bg = clrs[as.character(metadata$Site)], cex = 1.5)

mtext(side = 1, line = 2, text = "NMDS1")
mtext(side = 2, line = 3, text = "NMDS2" )

ordiellipse(trophicNMDS, metadata$Site == "BL", show.groups = TRUE, kind = "sd", col = "blue", lty = 1, lwd = 2.0)
ordiellipse(trophicNMDS, metadata$Site == "CB", show.groups = TRUE, kind = "sd", col = "red", lty = 1, lwd = 2.0)
ordiellipse(trophicNMDS, metadata$Site == "CH", show.groups = TRUE, kind = "sd", col = "dark green", lty = 1, lwd = 2.0)
ordiellipse(trophicNMDS, metadata$Site == "CM", show.groups = TRUE, kind = "sd", col = "coral", lty = 1, lwd = 2.0)
ordiellipse(trophicNMDS, metadata$Site == "CR", show.groups = TRUE, kind = "sd", col = "black", lty = 1, lwd = 2.0)
ordiellipse(trophicNMDS, metadata$Site == "Rt2", show.groups = TRUE, kind = "sd", col = "gray", lty = 1, lwd = 2.0)

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("topright", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))
```

I performed Kruskel-Wallace tests to see if any individual trophic modes were significantly different by site. None were. 

#### Guild

```{r guild setup, echo = F, include = F}

#need to know the unique trophic modes 
guild_names <- unique(guild_data$Guild)
# need somwhere to put data - a new shared file
n_guilds <- length(guild_names)
n_samples <- nrow(trans_matched_guild_shared)
all_guild_shared <- data.frame(matrix(0, nrow = n_samples, ncol = n_guilds))
rownames(all_guild_shared) <- rownames(trans_matched_guild_shared)
colnames(all_guild_shared) <- guild_names

#my loop
for(i in guild_names){
  guild_otus <- guild_data[guild_data$Guild == i, "OTU"]
  new_shared <- shared[ , guild_otus]
  if(length(guild_otus) > 1){
    guild_count <- apply(new_shared, 1, sum)
  } else {
    guild_count <- new_shared
  }
  all_guild_shared[ , i] <- guild_count
  
}

# 3. calculate the relative abundance
n_seqs <- apply(all_guild_shared, 1, sum)
guild_rel_abund <- all_guild_shared / n_seqs

tot_guild_rel_abund <- numeric()
for(h in colnames(guild_rel_abund)){
  y <- mean(guild_rel_abund[ ,h])
  tot_guild_rel_abund[h] <- y
}
sum(tot_guild_rel_abund)
tot_guild_rel_abund

L <- sort.list(tot_guild_rel_abund, decreasing = T)
tot_guild_rel_abund <- tot_guild_rel_abund[L]

```

Here are the top 10 functional guilds by relative abundance. The bar plot is numbered by the most abundant to least. Refer to the list for guild names

```{r guild plot, echo = F, include=T}
tot_guild_rel_abund[1:10]
guild_num_names <- tot_guild_rel_abund
names(guild_num_names) <- c(1:24)
barplot(height = guild_num_names[1:10], beside=T)

```

```{r per by guild, echo =F, include = T}
perm_guild <- adonis(all_guild_shared ~ Site * Lineage, data = metadata)
perm_guild
```
The permanova by functional guild still shows significant difference by site


I performed Kruskel-Wallace tests to see if any individual guilds were significantly different by site. None were. 

## Nutrients
### Soil Nutrients
```{r nutrient setup, echo = F, include=F, cache=T}
library(dplyr)
#importing leco file that measured C and N from 15 & 16 soils, and 16 plants
leco_dec16 <- read.table(file="~/git_repos/site_nutrients/LECO/Bickford_Dec16.csv", sep=",", header=F)
#importing leco file that measured C and N from some 15 soils
leco_mar16 <- read.table(file = "~/git_repos/site_nutrients/LECO/soils/LECO_soils_15.csv", sep = ",", header = T, stringsAsFactors = F)

head(leco_dec16)
head(leco_mar16)
colnames(leco_dec16) <- c("Project", "Sample", "Mass", "Percent_N", "Percent_C", "Time")
head(leco_dec16)

colnames(leco_mar16) <- c("Sample", "Mass", "Percent_N", "Percent_C")
head(leco_mar16)

myindex <- grep("15_Soil", leco_dec16$Project)
myindex

Soils15 <- leco_dec16[myindex, ]
head(Soils15)

#remove subsample number from sample ID
Sample <- gsub(pattern = "\\-.*", replacement = "", x = Soils15$Sample)
Soils15$Sample <- Sample
Sample <- gsub(pattern = "\\_\\d", replacement = "", x = leco_mar16$Sample)
leco_mar16$Sample <- Sample

head(Soils15)
head(leco_mar16)

leco_mar16$Sample[41] <- "PLB_Inv1"
leco_mar16$Sample[42] <- "PLB_Inv1"
leco_mar16$Sample[43] <- "PLB_Nat1"

#combine data sets

Sample <- c(Soils15$Sample, leco_mar16$Sample)
Mass <- c(Soils15$Mass, leco_mar16$Mass)
Percent_N <- c(Soils15$Percent_N, leco_mar16$Percent_N)
Percent_C <- c(Soils15$Percent_C, leco_mar16$Percent_C)

Soils15 <- data.frame(Sample = Sample, Mass = Mass, Percent_N = Percent_N, Percent_C = Percent_C, stringsAsFactors = F)

#reorder in alphabetical order
L <- order(Soils15$Sample)
Soils15 <- Soils15[L, ]


#take the mean of the subsamples run in dec 16
N <- tapply(Soils15$Percent_N, Soils15$Sample, FUN = mean)
C <- tapply(Soils15$Percent_C, Soils15$Sample, FUN = mean)
mass <- tapply(Soils15$Mass, Soils15$Sample, FUN = mean)

#re-order the samples in alphabetical order because tapply outputs order them that way
Sample <- unique(Soils15$Sample)
L <- order(Sample)
Sample <- Sample[L]

Site <- gsub(pattern = "\\_.*", replacement = "", x = Sample)
Lin_rep <- gsub(pattern = ".*_", replacement = "", x = Sample)
Lineage <- gsub(pattern = "\\d", replacement = "", x = Lin_rep)
rep <- gsub(pattern = "Nat|Inv", replacement = "", x = Lin_rep)
Sample <- gsub(pattern = "\\_", replacement = "", x = Sample)

Soils15 <- data.frame(Sample = Sample, Site, Lineage, Rep = rep, Mass = mass, Percent_N = N, Percent_C = C, stringsAsFactors = F)
head(Soils15)

sample_meta <- c(Soils15$Sample, "BLInv4", "BLInv5","BLInv6","BLNat4", "BLNat5","BLNat6")
site_meta <- c(Soils15$Site, "BL","BL","BL","BL","BL","BL")
lin_meta <- c(Soils15$Lineage, "Inv","Inv","Inv","Nat","Nat","Nat")
rep_meta <- c(Soils15$Rep, 4,5,6,4,5,6)
mass_meta <- c(Soils15$Mass, NA,NA,NA,NA,NA,NA)
N_meta <- c(Soils15$Percent_N, NA,NA,NA,NA,NA,NA)
C_meta <- c(Soils15$Percent_C, NA,NA,NA,NA,NA,NA)

soils15_wnulls <- data.frame(Sample = sample_meta, Site = site_meta, 
                             Lineage = lin_meta, Rep = rep_meta, 
                             Mass = mass_meta, Percent_N = N_meta, 
                             Percent_C = C_meta, stringsAsFactors = F)

sample_meta <- c(soils15_wnulls$Sample, "CHInv4", "CHInv5","CHInv6","CHNat4", "CHNat5","CHNat6","Rt2Nat1")
site_meta <- c(soils15_wnulls$Site, "CH","CH","CH","CH","CH","CH", "Rt2")
lin_meta <- c(soils15_wnulls$Lineage, "Inv","Inv","Inv","Nat","Nat","Nat","Nat")
rep_meta <- c(soils15_wnulls$Rep, 4,5,6,4,5,6,1)
mass_meta <- c(soils15_wnulls$Mass, NA,NA,NA,NA,NA,NA,NA)
N_meta <- c(soils15_wnulls$Percent_N, NA,NA,NA,NA,NA,NA,NA)
C_meta <- c(soils15_wnulls$Percent_C, NA,NA,NA,NA,NA,NA,NA)



soils15_wnulls <- data.frame(Sample = sample_meta, Site = site_meta, 
                             Lineage = lin_meta, Rep = rep_meta, 
                             Mass = mass_meta, Percent_N = N_meta, 
                             Percent_C = C_meta, stringsAsFactors = F)



o <- order(soils15_wnulls$Sample)
soils15_wnulls <- soils15_wnulls[o, ]

rownames(soils15_wnulls) <- soils15_wnulls$Sample

source("~/git_repos/Root_ITS/data_setup.R")
soils_only_in_shared <- soils15_wnulls[rownames(shared), ]
soils_only_in_shared <- soils_only_in_shared[(rownames(soils_only_in_shared)) != "NA.1", ]

meta_soils <- soils_only_in_shared
```
```{r plot of nitrogen, echo = F, include = T}

clrs2 <- c(Inv = "blue", Nat = "red")
meta_soils_finite <- meta_soils[is.finite(meta_soils$Percent_N), ]
meta_soils_finite$Lineage <- as.factor(meta_soils_finite$Lineage)
meta_soils_finite$Site <- as.factor(meta_soils_finite$Site)
meta_soils_finite$Percent_N <- as.numeric(meta_soils_finite$Percent_N)

stripchart(Percent_N ~ Lineage + Site, data = meta_soils_finite, vertical = T, ylab = "Percent Soil N", ylim = c(0,3), xlab = NULL, axes = F, pch = 20, jitter = 0.1, col = clrs2[levels(meta_soils_finite$Lineage)])

mtext(side=1, at=c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), line=1, text = c("BL", "CB", "CH", "CM", "CR", "PLB", "Rt2", "SB"), cex = 1, font = 1)

axis(1, at = c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), labels = F)
axis(2, labels = T)

legend_text <- c("Non-Native", "Native")
legend("topright", legend_text, pch = 15,
       col = c("blue","red"))

```

```{r boxplot of carbon, echo = F, include = T}
clrs2 <- c(Inv = "blue", Nat = "red")
meta_soils_finite <- meta_soils[is.finite(meta_soils$Percent_C), ]
meta_soils_finite$Lineage <- as.factor(meta_soils_finite$Lineage)
meta_soils_finite$Site <- as.factor(meta_soils_finite$Site)
meta_soils_finite$Percent_C <- as.numeric(meta_soils_finite$Percent_C)

stripchart(Percent_C ~ Lineage + Site, data = meta_soils_finite, vertical = T, ylab = "Percent Soil C", ylim = c(0,40), xlab = NULL, axes = F, pch = 20, jitter = 0.1, col = clrs2[levels(meta_soils_finite$Lineage)])

mtext(side=1, at=c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), line=1, text = c("BL", "CB", "CH", "CM", "CR", "PLB", "Rt2", "SB"), cex = 1, font = 1)

axis(1, at = c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), labels = F)
axis(2, labels = T)

legend_text <- c("Non-Native", "Native")
legend("topright", legend_text, pch = 15,
       col = c("blue","red"))

```

plot NMDS scores against nitrogen content

```{r nutrients by community, echo = F, include = T}

reg_n <- lm(meta_soils$Percent_N ~ NMDS1 * NMDS2)
reg_n
summary(reg_n)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1.5,1.5), ylim = c(0,3))
axis(side = 1)
axis(side = 2, las = 1)
points(x = NMDS_scores$NMDS1, y = meta_soils$Percent_N, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1)

mtext(side = 1, line = 2, text = "NMDS1 scores")
mtext(side = 2, line = 3, text = "Percent N" )

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("topright", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```

Now plot against total C

```{r Carbon by community, echo = F, include = T}

reg_c <- lm(meta_soils$Percent_C ~ NMDS1 * NMDS2)
reg_c
summary(reg_c)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1.5,1.5), ylim = c(0,40))
axis(side = 1)
axis(side = 2, las = 1)
points(x = NMDS_scores$NMDS1, y = meta_soils$Percent_C, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1)

mtext(side = 1, line = 2, text = "NMDS1 scores")
mtext(side = 2, line = 3, text = "Percent C" )

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("topright", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```

Adding nutrients into PerMANOVA

```{r perm with nutrients, echo = F, include = T}
meta_soils$Percent_N <- as.numeric(meta_soils$Percent_N)
meta_soils$Percent_C <- as.numeric(meta_soils$Percent_C)

meta_soils_finite <- meta_soils[is.finite(meta_soils$Percent_N), ]
shared_finite <- shared[rownames(meta_soils_finite) , ]

perm_nutrient <- adonis(shared_finite ~ Site + Lineage + Percent_C + Percent_N, data = meta_soils_finite, method = "bray")
perm_nutrient
```

plot trophic NMDS scores against N content

```{r regression trophic mode (N), echo = F, include = T}
trophicreg_n <- lm(meta_soils$Percent_N ~ trophic_scores$NMDS1 * trophic_scores$NMDS2)
trophicreg_n
summary(trophicreg_n)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1.5, 1), ylim = c(0,3))
axis(side = 1)
axis(side = 2, las = 1)
points(x = trophic_scores$NMDS2, y = meta_soils$Percent_N, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1)

mtext(side = 1, line = 2, text = "NMDS2 scores")
mtext(side = 2, line = 3, text = "Percent N" )

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("left", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```

plot percent C against trophic mode scores. Used axis 2 here because it was closer to significant than axis 1

```{r regression trophic mode (C), echo = F, include = T}
trophicreg_c <- lm(meta_soils$Percent_C ~ trophic_scores$NMDS1 * trophic_scores$NMDS2)
trophicreg_c
summary(trophicreg_c)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1.5,1), ylim = c(0,40))
axis(side = 1)
axis(side = 2, las = 1)
points(x = trophic_scores$NMDS2, y = meta_soils$Percent_C, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1)

mtext(side = 1, line = 2, text = "NMDS2 scores")
mtext(side = 2, line = 3, text = "Percent C" )

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("left", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```
### Tissue Nutrients

```{r tissue nutrient setup, echo = F, include=F, cache=T}
tissue_15 <- read.table("~/git_repos/site_nutrients/LECO/tissue/leaf/leco_7_28_withcodes.csv", header = F, sep = ",", stringsAsFactors = F)
tissue_15 <- tissue_15[ , -c(10:13)]
tissue_15 <- tissue_15[ , -c(3,4,6)]

colnames(tissue_15) <- c("Project", "Code", "Sample", "Mass", "Percent_N", "Percent_C")
tissue_15 <- tissue_15[tissue_15$Project == "Bickford_2016", ]
tissue_15 <- tissue_15[is.finite(tissue_15$Percent_N), ]

N <- tapply(tissue_15$Percent_N, tissue_15$Sample, FUN = mean)
C <- tapply(tissue_15$Percent_C, tissue_15$Sample, FUN = mean)
mass <- tapply(tissue_15$Mass, tissue_15$Sample, FUN = mean)

Sample <- unique(tissue_15$Sample)
L <- order(Sample)
Sample <- Sample[L]

Site <- gsub(pattern = "Nat.|Inv.", replacement = "", x = Sample)
Lin_rep <- gsub(pattern = "BL|CB|CH|CM|CR|PLB|Rt2|SB", replacement = "", x = Sample)
Lineage <- gsub(pattern = "\\d", replacement = "", x = Lin_rep)
rep <- gsub(pattern = "Nat|Inv", replacement = "", x = Lin_rep)

tissue_15 <- data.frame(Sample = Sample, Site, Lineage, Rep = rep, Mass = mass, Percent_N = N, Percent_C = C, stringsAsFactors = F)
head(tissue_15)

sample_meta <- c(tissue_15$Sample, "BLInv2")
site_meta <- c(tissue_15$Site, "BL")
lin_meta <- c(tissue_15$Lineage, "Inv")
rep_meta <- c(tissue_15$Rep, 2)
mass_meta <- c(tissue_15$Mass, NA)
N_meta <- c(tissue_15$Percent_N, NA)
C_meta <- c(tissue_15$Percent_C, NA)

tissue15_wnulls <- data.frame(Sample = sample_meta, Site = site_meta, 
                             Lineage = lin_meta, Rep = rep_meta, 
                             Mass = mass_meta, Percent_N = N_meta, 
                             Percent_C = C_meta, stringsAsFactors = F)

o <- order(tissue15_wnulls$Sample)
tissue15_wnulls <- tissue15_wnulls[o, ]

rownames(tissue15_wnulls) <- tissue15_wnulls$Sample

source("~/git_repos/Root_ITS/data_setup.R")
source("~/git_repos/Root_ITS/community analysis.R")
tissue_only_in_shared <- tissue15_wnulls[rownames(shared), ]

meta_tissue <- tissue_only_in_shared
```

```{r plot of tissue nitrogen, echo = F, include = T}

clrs2 <- c(Inv = "blue", Nat = "red")
meta_tissue_finite <- meta_tissue[is.finite(meta_tissue$Percent_N), ]
meta_tissue_finite$Lineage <- as.factor(meta_tissue_finite$Lineage)
meta_tissue_finite$Site <- as.factor(meta_tissue_finite$Site)
meta_tissue_finite$Percent_N <- as.numeric(meta_tissue_finite$Percent_N)

stripchart(Percent_N ~ Lineage + Site, data = meta_tissue_finite, vertical = T, ylab = "Percent Tissue N", ylim = c(0,4), xlab = NULL, axes = F, pch = 20, jitter = 0.1, col = clrs2[levels(meta_tissue_finite$Lineage)])

mtext(side=1, at=c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), line=1, text = c("BL", "CB", "CH", "CM", "CR", "PLB", "Rt2", "SB"), cex = 1, font = 1)

axis(1, at = c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5), labels = F)
axis(2, labels = T)

legend_text <- c("Non-Native", "Native")
legend("topright", legend_text, pch = 15,
       col = c("blue","red"))

```

```{r tissue nutrients by community, echo = F, include = T}

reg_tis_n <- lm(meta_tissue$Percent_N ~ NMDS1 * NMDS2)
reg_tis_n
summary(reg_tis_n)

clrs <- c(BL = "blue", CB = "red", CH = "dark green", CM = "coral", CR = "black", PLB = "orange", Rt2 = "gray", SB = "purple")
pchs <- c(Nat = 21, Inv = 23)

quartz.options(height=5, width=5)
plot.new() 
par(oma = c(1, 1, 1, 1))
par(mar = c( 5, 5, 0, 0 ))
plot.window(xlim = c(-1.5,1.5), ylim = c(0,4))
axis(side = 1)
axis(side = 2, las = 1)
points(x = NMDS_scores$NMDS1, y = meta_tissue$Percent_N, pch = pchs[as.character(metadata$Lineage)], 
      bg = clrs[as.character(metadata$Site)], cex = 1)

mtext(side = 1, line = 2, text = "NMDS1 scores")
mtext(side = 2, line = 3, text = "Percent N" )

legend_text <- c("Bullard Lake", "Cecil Bay", "Chelsea", "Cheboygan Marsh", "Castle Rock", "Point LeBarb", "Route 2", "Sturgeon Bay")
legend("topright", legend_text, pch = 15, 
       col = c("blue","red","dark green","coral","black","orange","gray","purple"))

legend_text2 <- c("Native", "Non-Native")
legend("topleft", legend_text2, pch = c(1,5))


```